---
title: "R Notebook"
output: html_notebook
---

2.  Small Dataset - *US Violent Crime*

Dataset info: <https://www.kaggle.com/datasets/mathchi/violent-crime-rates-by-us-state>

Size: 1.35KB (50 x 5)

```{r}
# Library imports

library(cluster)
library(factoextra)
library(utils)
library(ClusterR)
library(microbenchmark)

```

```{r}
# Loading the dataset and preview of the data

data1 <- read.csv('Datasets/US_violent_crime.csv')

dim(data1)

head(data1, 10)
```

```{r}
# Find optimal number of clusters

fviz_nbclust(data1, hcut, method = "silhouette")
```

```{r}
# Apply CLARA

# Variable parameters
clusterNr <- 2
samples <- 25
sampSize <- 5

cl.res <- clara(data1, clusterNr, metric = 'manhattan', samples = samples, sampsize = sampSize, pamLike = TRUE)
```

```{r}
# Plot of CLARA clustering on 'US violent crime'

fviz_cluster(cl.res, data1, geom = "point", repel = TRUE, show.clust.cent = TRUE, ellipse.type = "norm", palette = "Set2", ggtheme = theme_minimal(), axes = c(1,5))
```

```{r}
# Trying out the Clara_Medoids function from ClusterR library

#df <- as.data.frame(lapply(data1, as.integer))
#nrow(df[!apply(df == "?", 1, all),])

df = subset(data1, select = -c(X) )

clR <- Clara_Medoids(df, clusters = clusterNr, distance_metric = 'manhattan', 
  samples = samples, sample_size = 0.2, swap_phase = TRUE, verbose = F, threads = 1)

fviz_cluster(cl.res, data1, geom = "point", ellipse.type = "norm", main = "Clara (ClusterR)")
```

```{r}
# Kmeans plot of same data

km.res <- kmeans(df, clusterNr)

fviz_cluster(km.res, df, geom = "point", ellipse.type = "norm", main = "Kmeans")
```

```{r}
# Measure the running time using microbenchmark

result <- microbenchmark(
  clara(data1, clusterNr, metric = 'manhattan', samples = samples, sampsize = sampSize, pamLike = TRUE),
  kmeans(df, clusterNr),
  times = 100
)

print(result)
boxplot(result, names=c('clara', 'kmeans'))
```

[Conclusion:]{.underline}

-   Regarding the actual clusters, the differences are not huge but we can observe that *clara* is better at grouping the points near the border than *kmeans*.

-   There isn't a big difference with execution times between *clara* and *kmeans* when the data is not so large, *clara* being more inefficient on this task. The difference can be seen once the dataset increases in size.
